---
title: "Exploratory Analyses"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, message = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(plotly)
library(patchwork)
library(ggpubr)
library(ggplot2)
```


In Exploratory Analyses, we are focusing on product count, unit price, revenue, and time trend by using pie, donut, bar, line, and scatter plot.

<p>&nbsp;</p >
## Summary of packages
tidyverse, lubridate, plotly, patchwork, ggpubr, ggplot2

```{r,data cleaning, message = FALSE}
bakery_df =
  read_csv("./Data/Bakery_sales.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    unit_price = str_replace(unit_price, "€", ""),
    unit_price = str_replace(unit_price, ",", "."),
    unit_price = as.numeric(unit_price),
    product_name = article) %>% 
  filter(product_name != ".") %>% 
  select(-article)
```

## Introduction & Background 

We used updated version dataset of a French bakery from 2021-01-01 to 2022-09-30. The dataset provides the daily transaction details of customers. Since the dataset is still updating, we used downloaded version on 2022-12-03.

The dataset includes:

`date`: date order

`time`: time order

`ticket number`: identifier for every single transaction

`article`: name of the product sold (in French)

`quantity`: quantity sold

`unit_price`: price per product (in Euro)

`objective`: forecast the sales in order to ease the production planning

It contains `r ncol(bakery_df)` variables and `r nrow(bakery_df)` observations.

<p>&nbsp;</p >
Firstly, we used `read_csv()` to import the dataset "Bakery_sales.csv", and applied `janitor::clean_names()` to change the variables' names into lower cases. Next, we tidied dataset by applying `mutate()` to remove the euro sign "€", change "," to ".", convert `unit_price` to numeric form, and renamed the `article` as `product_name`. Finally, we deleted the missing data via `filter()` and `article` via `select()`. 


<p>&nbsp;</p >
## Visualization

### Top 10 favorite bakery products
```{r}
plot_pie =
  bakery_df %>% 
  group_by(product_name) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs)) %>% 
  head(10) %>% 
  plot_ly(labels = ~product_name, values = ~n_obs, type = 'pie', colors = "viridis") %>% 
  layout(xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

plot_pie
```

The pie chart shows the total of sale counts of top 10 favorite bakery products from 2021-01-01 to 2022-09-30. As the plot shows that the `TRADITIONAL BAGUETTE` (67,689; 42.5%) is the most favorite bakery product. The rest ranking from top 2 to top 10 favorite bakeries are  `COUPE` (20,470; 12.90%), `BAGUETTE` (15,292; 9.61%) `BANETTE` (15,130; 9.51%), `CROISSANT` (11,508; 7.23%), `PAIN AU CHOCOLAT` (10,578; 6.65%), `SPECIAL BREAD` (5195; 3.26%), `CEREAL BAGUETTE` (4961; 3.12%), `FORMULE SANDWITCH` (4202; 2.64%), and `BOULE 400G` (4099; 2.58%), respectively. 

<p>&nbsp;</p >
### Top 10 favorite bakery products in 2021/2022
```{r}
plot_pie =
  bakery_df %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.numeric(year),
         month = as.integer(month),
         day = as.integer(day),) 
  
plot_pie_2021 = 
  plot_pie %>% 
  filter(year == 2021) %>%
  group_by(product_name) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs)) %>% 
  head(10) %>% 
  plot_ly(labels = ~product_name, values = ~n_obs) %>% 
  add_pie(hole = 0.6) %>% 
  layout(title = 'Top 10 favorite bakery products in 2021',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

plot_pie_2022 = 
  plot_pie %>% 
  filter(year == 2022) %>%
  group_by(product_name) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs)) %>% 
  head(10) %>% 
  plot_ly(labels = ~product_name, values = ~n_obs) %>% 
  add_pie(hole = 0.6) %>% 
  layout(title = 'Top 10 favorite bakery products in 2022',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

plot_pie_2021
plot_pie_2022
```

The first donut chart shows the total of sale counts of top 10 favorite bakery products from 2021-01-01 to 2021-12-31. As the plot shows that the `TRADITIONAL BAGUETTE` (37,399; 42%) is the most favorite bakery . The rest ranking from top 2 to top 10 favorite bakeries are  `COUPE` (11,759; 13.2%), `BAGUETTE` (8713; 9.78%) `BANETTE` (8575; 9.63%), `CROISSANT` (6291; 7.06%), `PAIN AU CHOCOLAT` (5739; 6.44%), `CEREAL BAGUETTE` (3070; 3.45%), `SPECIAL BREAD` (2936; 3.3%), `CAMPAGNE` (2289; 2.57%), and `BOULE 400G` (2283; 2.46%), respectively. 

The second donut chart shows the total of sale counts of top 10 favorite bakery products from 2022-01-01 to 2022-09-30. As the plot shows that the `TRADITIONAL BAGUETTE` (30,290; 43.2%) is the most favorite bakery . The rest ranking from top 2 to top 10 favorite bakeries are  `COUPE` (8711; 12.4%), `BAGUETTE` (6579; 9.38%) `BANETTE` (6555; 9.35%), `CROISSANT` (5217; 7.44%), `PAIN AU CHOCOLAT` (4839; 6.9%), `SPECIAL BREAD` (2259; 3.22%), `FORMULE SANDWITCH` (1976; 2.82%), `CEREAL BAGUETTE` (1891; 2.7%), and `BOULE 400G` (1816; 2.59%), respectively. 

Compared donut charts of 2021 and 2022, the ranking of top 1-6 favorite bakeries does not change. `CAMPAGNE` dropped out of the top 10 list, and `FORMULE SANDWITCH` was added the top 10 favorite bakery products in 2022. Top 7-9 moved slightly from 2021 to 2022. Top 10 `BOULE 400G` stayed in the same position. 

<p>&nbsp;</p >
### Top 10 favorite bread monthly sale counts in 2021
```{r}
top_line =
  bakery_df %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.numeric(year),
         month = as.integer(month),
         day = as.integer(day),) %>%
  filter(year == 2021) %>% 
  select (year, month, product_name, quantity) %>% 
  group_by (year, month, product_name) %>% 
  mutate (total_counts = sum (quantity)) %>% 
  distinct(year, month, product_name,total_counts) 

top_line1_2021 =
  top_line %>% 
  filter(product_name %in% c("TRADITIONAL BAGUETTE")) %>% 
  ggplot(aes(x =month, y = total_counts, color = product_name)) + 
    geom_line() +
    theme(axis.text.x = element_text(angle=90,hjust = 1), legend.position = "none") +
    labs(x = "month",
         y = "counts of TRADITIONAL BAGUETTE",
         title = "Top 1") +
    scale_x_continuous(
    breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12), 
    labels = c("Jan","Feb","Mar","Apr","May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) 

top_line2_2021 =
  top_line %>% 
    filter(product_name %in% c("COUPE", "BAGUETTE","BANETTE", "CROISSANT", "PAIN AU CHOCOLAT", "CEREAL BAGUETTE", "SPECIAL BREAD", "CAMPAGNE", "BOULE 400G")) %>% 
  ggplot(aes(x =month, y = total_counts, color = product_name)) + 
    geom_line() +
    theme(axis.text.x = element_text(angle=90,hjust = 1), legend.position = "none", legend.title = element_text(size=7), legend.text = element_text(size=7)) + 
    labs(x = "month",
         y = "counts of bread",
         title = "Top 2-10") +
    scale_x_continuous(
    breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12), 
    labels = c("Jan","Feb","Mar","Apr","May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

ggarrange(top_line2_2021, top_line1_2021, ncol=2, nrow=1, common.legend = TRUE, legend="bottom")
```

In general, monthly sale counts in 2021 increased dramatically from June to August.

For `TRADITIONAL BAGUETTE` (line plot on the right), total monthly sale counts in 2021 peaked at August (>11,250). However, total monthly sale counts decreased from May to June, August to September, and October to December in 2021. 

For line plot on the left, total monthly sale counts in 2021 are less than 3,000 but greater than 500 in every month for top 2-6 (`COUPE`, `BAGUETTE`, `BANETTE`, `CROISSANT`, `PAIN AU CHOCOLAT`). For top 7-10 (`CEREAL BAGUETTE`, `SPECIAL BREAD`, `CAMPAGNE`, and `BOULE 400G`), total monthly sale counts in 2021 are less than 1,000. 

<p>&nbsp;</p >
### Top 10 favorite bread monthly sale counts in 2022
```{r}
top_line =
  bakery_df %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.numeric(year),
         month = as.integer(month),
         day = as.integer(day),) %>%
  filter(year == 2022) %>% 
  select (year, month, product_name, quantity) %>% 
  group_by (year, month, product_name) %>% 
  mutate (total_counts = sum (quantity)) %>% 
  distinct(year, month, product_name,total_counts) 

top_line1_2022 =
  top_line %>% 
  filter(product_name %in% c("TRADITIONAL BAGUETTE")) %>% 
  ggplot(aes(x =month, y = total_counts, color = product_name)) + 
    geom_line() +
    theme(legend.position = "none", axis.text.x = element_text(angle=90,hjust = 1)) + 
    labs(x = "month",
         y = "counts of TRADITIONAL BAGUETTE",
         title = "Top 1") +
    scale_x_continuous(
    breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12), 
    labels = c("Jan","Feb","Mar","Apr","May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) 

top_line2_2022 =
  top_line %>% 
    filter(product_name %in% c("COUPE", "BAGUETTE","BANETTE", "CROISSANT", "PAIN AU CHOCOLAT", "SPECIAL BREAD", "FORMULE SANDWICH","CEREAL BAGUETTE", "BOULE 400G")) %>% 
  ggplot(aes(x =month, y = total_counts, color = product_name)) + 
    geom_line() +
    theme(axis.text.x = element_text(angle=90,hjust = 1), legend.position = "none", legend.title = element_text(size=7), legend.text = element_text(size=7)) + 
    labs(x = "month",
         y = "counts of bread",
         title = "Top 2-10") +
    scale_x_continuous(
    breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12), 
    labels = c("Jan","Feb","Mar","Apr","May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) 
 
ggarrange(top_line2_2022, top_line1_2022, ncol=2, nrow=1, common.legend = TRUE, legend="bottom")

```

In general, monthly sale counts in 2022 show similar trend as in 2021, but monthly sale counts end at 2022-09-30 due to limited data. 

For `TRADITIONAL BAGUETTE` (line plot on the right), total monthly sale counts in 2022 peaked at August (>12,500). However, total monthly sale counts decreased from May to June and August to September in 2022. Compared to 2021 (>11,250), total monthly sale counts of `TRADITIONAL BAGUETTE` is higher in 2022 (>12,500). 

For line plot on the left, total monthly sale counts in 2022 are less than 3,500 but greater than 500 in every month for top 2-6 (`COUPE`, `BAGUETTE`, `BANETTE`, `CROISSANT`, `PAIN AU CHOCOLAT`). For top 7-10 (`SPECIAL BREAD`, `FORMULE SANDWICH`, `CEREAL BAGUETTE`, and `BOULE 400G`), total monthly sale counts in 2022 are less than 600. 

<p>&nbsp;</p >
### Ranking of mean prices of top 10 products in 2021
```{r}
mean_price_2021 =
  bakery_df %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.numeric(year),
         month = as.integer(month),
         day = as.integer(day)) %>%
  filter(year == 2021) %>% 
  filter(product_name %in% c("TRADITIONAL BAGUETTE", "COUPE", "BAGUETTE","BANETTE", "CROISSANT", "PAIN AU CHOCOLAT", "CEREAL BAGUETTE", "SPECIAL BREAD", "CAMPAGNE", "BOULE 400G")) %>% 
  group_by(product_name) %>% 
  summarize(mean_price = mean(unit_price)) %>% 
  arrange(desc(mean_price)) %>% 
  mutate(product_name = forcats::fct_relevel(product_name, c("TRADITIONAL BAGUETTE", "COUPE", "BAGUETTE","BANETTE", "CROISSANT", "PAIN AU CHOCOLAT", "CEREAL BAGUETTE", "SPECIAL BREAD", "CAMPAGNE", "BOULE 400G"))) %>% 
  ggplot(aes(x =product_name, y = mean_price, color = product_name)) + 
      geom_point() +
      geom_text(aes(label=mean_price), vjust = 0, nudge_y = 0.1) +
      ylim(0,3) +
    theme(axis.text.x = element_text(angle=60,hjust = 1), legend.position = "none") + 
    labs(x = "name of products",
         y = "mean prices in euro") 
  
mean_price_2021 
```

The scatter plot shows the ranking of mean prices of top 10 favorite bakery products in 2021, with most popularity to least popularity (top1-10) from the left to the right on x-axis.

`SPECIAL BREAD` has the highest mean price (2.4 €) while `COUPE` has the lowest mean price (0.15 €). The mean prices of `TRADITIONAL BAGUETTE` and 
`PAIN AU CHOCOLAT` are the same (1.2 €).

<p>&nbsp;</p >
### Ranking of mean prices of top 10 products in 2022
```{r, warning=FALSE}
mean_price_2022 =
  bakery_df %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.numeric(year),
         month = as.integer(month),
         day = as.integer(day)) %>%
  filter(year == 2022) %>% 
  filter(product_name %in% c("TRADITIONAL BAGUETTE", "COUPE", "BAGUETTE","BANETTE", "CROISSANT", "PAIN AU CHOCOLAT", "SPECIAL BREAD", "FORMULE SANDWICH","CEREAL BAGUETTE", "BOULE 400G")) %>% 
  group_by(product_name) %>% 
  summarize(mean_price = mean(unit_price)) %>%
  mutate(across(where(is.numeric), ~ round(., digits = 2))) %>% 
  arrange(desc(mean_price)) %>% 
  mutate(product_name = forcats::fct_relevel(product_name, c("TRADITIONAL BAGUETTE", "COUPE", "BAGUETTE","BANETTE", "CROISSANT", "PAIN AU CHOCOLAT", "SPECIAL BREAD", "FORMULE SANDWICH","CEREAL BAGUETTE", "BOULE 400G"))) %>% 
  ggplot(aes(x =product_name, y = mean_price, color = product_name)) + 
      geom_point() +
      geom_text(aes(label=mean_price), vjust = 0, nudge_y = 0.2) +
      ylim(0,7.5) +
    theme(axis.text.x = element_text(angle=60,hjust = 1), legend.position = "none") + 
    labs(x = "name of products",
         y = "mean prices in euro") 
  
mean_price_2022 
```

The scatter plot shows the ranking of mean prices of top 10 favorite bakery products in 2022, with most popularity to least popularity (top1-10) from the left to the right on x-axis. In general, the mean price of `COUPE` stayed the same from 2021 to 2022, but mean prices of other breads increased slightly.

`FORMULE SANDWICH` has the highest mean price (6.94 €) while `COUPE` has the lowest mean price (0.15 €). The mean prices of `TRADITIONAL BAGUETTE` and 
`PAIN AU CHOCOLAT` are the same (1.27 €)., but both increased from 1.2 € to 1.27 € when compared to the scatter plot in 2021. 

<p>&nbsp;</p >
### Ranking of total revenue of top 10 products in 2021 and 2022
```{r}
total_rev_bar=
  bakery_df %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.numeric(year),
         month = as.integer(month),
         day = as.integer(day)) %>%
  group_by(year, product_name) %>% 
  mutate(rev = quantity * unit_price) %>% 
  summarize(prod_rev = sum(rev)) %>% 
  arrange(desc(prod_rev)) 
  
total_rev_bar_2021=
  total_rev_bar %>% 
  filter(year == 2021) %>% 
  filter(product_name %in% c("TRADITIONAL BAGUETTE", "COUPE", "BAGUETTE","BANETTE", "CROISSANT", "PAIN AU CHOCOLAT", "CEREAL BAGUETTE", "SPECIAL BREAD", "CAMPAGNE", "BOULE 400G")) %>% 
  mutate(product_name = fct_reorder(product_name, prod_rev)) %>% 
  mutate(across(where(is.numeric), ~ round(., digits = 2))) %>% 
  ggplot(aes(x =product_name, y= prod_rev, fill=product_name)) + 
    geom_bar(stat="identity", width=0.5) +
    geom_text(aes(label=prod_rev), vjust = 0, nudge_y = 100, size=3) +
    ylim(0,80000) +
    labs(x = "name of products",
         y = "total revenue in euro",
         title = "ranking in 2021") +
  theme(axis.text.x = element_text(angle = 15, hjust = 1), legend.position = "none")
  
total_rev_bar_2022=
  total_rev_bar %>% 
  filter(year == 2022) %>% 
  filter(product_name %in% c("TRADITIONAL BAGUETTE", "COUPE", "BAGUETTE","BANETTE", "CROISSANT", "PAIN AU CHOCOLAT", "SPECIAL BREAD", "FORMULE SANDWICH","CEREAL BAGUETTE", "BOULE 400G")) %>% 
  mutate(product_name = fct_reorder(product_name, prod_rev)) %>%
  mutate(across(where(is.numeric), ~ round(., digits = 2))) %>% 
  ggplot(aes(x =product_name, y= prod_rev, fill=product_name)) + 
    geom_bar(stat="identity", width=0.5) +
    geom_text(aes(label=prod_rev), vjust = 0, nudge_y = 100, size=3) +
    ylim(0,80000) +
    labs(x = "name of products",
         y = "total revenue in euro",
         title = "ranking in 2022") +
  theme(axis.text.x = element_text(angle = 15, hjust = 1), legend.position = "none")
 
ggarrange(total_rev_bar_2021, total_rev_bar_2022,  heights = c(5, 5),
          ncol = 1, nrow = 2)
```

The bar chart shows the ranking of total revenue of top 10 products in 2021 and 2022. The total revenues of `TRADITIONAL BAGUETTE` are the highest (78742.8 € in 2021; 66013.25 € in 2022) while the total revenues of `COUPE` are the lowest (2049.9 € in 2021; 1475.85 € in 2022) in the two years respectively. The total revenues of `TRADITIONAL BAGUETTE` is approximately 38 times the total revenues of `COUPE` in 2021 and 45 times of that in 2022.

<p>&nbsp;</p >
### Distribution of monthly revenue in 2021 and 2022
```{r}
dis_rev_bar=
  bakery_df %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.numeric(year),
         month = as.integer(month),
         day = as.integer(day),) 

dis_rev_bar_2021=
  dis_rev_bar %>%
    filter(year == 2021,
         product_name %in% c("TRADITIONAL BAGUETTE", "COUPE", "BAGUETTE","BANETTE", "CROISSANT", "PAIN AU CHOCOLAT", "CEREAL BAGUETTE", "SPECIAL BREAD", "CAMPAGNE", "BOULE 400G")) %>% 
  group_by(year, month, product_name) %>% 
  mutate(rev = quantity * unit_price) %>% 
  summarize(prod_dis_rev = sum(rev)) %>% 
  arrange(desc(prod_dis_rev)) %>% 
   mutate(product_name = fct_reorder(product_name, prod_dis_rev)) %>% 
    ggplot(aes(x =month, y= prod_dis_rev, fill=product_name)) + 
    scale_fill_viridis_d()+
    geom_bar(stat="identity", position="stack") +
    labs(x = "month",
         y = "total revenue in euro",
         title = "Distribution of monthly revenue in 2021") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "right", legend.title = element_text(size=7), legend.text = element_text(size=7), legend.key.size = unit(0.2, 'cm')) +
  scale_x_continuous(
    breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12), 
    labels = c("Jan","Feb","Mar","Apr","May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) 

dis_rev_bar_2022=
  dis_rev_bar %>%
    filter(year == 2022,
         product_name %in% c("TRADITIONAL BAGUETTE", "COUPE", "BAGUETTE","BANETTE", "CROISSANT", "PAIN AU CHOCOLAT", "SPECIAL BREAD", "FORMULE SANDWICH","CEREAL BAGUETTE", "BOULE 400G")) %>% 
  group_by(year, month, product_name) %>% 
  mutate(rev = quantity * unit_price) %>% 
  summarize(prod_dis_rev = sum(rev)) %>% 
  arrange(desc(prod_dis_rev)) %>% 
   mutate(product_name = fct_reorder(product_name, prod_dis_rev)) %>% 
    ggplot(aes(x =month, y= prod_dis_rev, fill=product_name)) + 
    scale_fill_viridis_d()+
    geom_bar(stat="identity", position="stack") +
    labs(x = "month",
         y = "total revenue in euro",
         title = "Distribution of monthly revenue in 2022") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "right",legend.title = element_text(size=7), legend.text = element_text(size=7), legend.key.size = unit(0.2, 'cm')) +
  scale_x_continuous(
    breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12), 
    labels = c("Jan","Feb","Mar","Apr","May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) 

ggarrange(dis_rev_bar_2021, dis_rev_bar_2022,  heights = c(5, 5),
          ncol = 1, nrow = 2)
```

The bar plots show distributions of monthly revenue in 2021 and 2022. `TRADITIONAL BAGUETTE` has the largest proportion of monthly revenues (around 50%) in 2021 and 2022. Moreover, the revenue trend from Jan to Dec is similar in 2021 and 2022. Since the bakery sold more products in summer than winter, and reached peak in August, we could expect higher revenue in summer than winter, and the highest revenue in August. 


<p>&nbsp;</p >
#### Summary 
`TRADITIONAL BAGUETTE` is the most favorite bread in this French bakery store. French people LOVE TRADITIONAL BAGUETTE. 



